---
title: "Description of the IBI/HR data"
author: "Jussi Korpela"
date: '2018-08-08'
output:
  html_document:
    theme: united
    toc: yes
  github_document: default
  pdf_document:
    toc: yes
params:
  #input_data_rds: /home/jussi/Dropbox/Shared/CORE_jkor/ukonkuva/analysis/HRV_blocks_noldus_ajo2alk_cw600/ibi/stancon18_fits/ibi-data_train4stan.rds
  input_data_rds: /home/jkor/Dropbox/Shared/CORE_jkor/ukonkuva/analysis/HRV_blocks_noldus_ajo2alk_cw600/ibi/stancon18_fits/ibi-syndata_train4stan.rds
  #opts: OPTS
---

# Info

input dir: `r dirname(params$input_data_rds)`

input file name: `r basename(params$input_data_rds)`

# Rationale

Provides descriptive statistics of the IBI/HR dataset to help set informative priors and thus make models identify correctly.


# Setup
```{r setup, include=FALSE}
require(tidyverse)
require(ggplot2)
require(cowplot)

# Set general options
knitr::opts_knit$set(echo = TRUE,
                     fig.width = 6,
                     fig.asp = 0.618,
                     out.width = "100%",
                     fig.align = "center")
```


```{r}
# A helper for saving
my_saveplot <- function(p, savename,
                        savedir = file.path('/home/jkor/Dropbox/Shared/CORE_jkor/ukonkuva/analysis/HRV_blocks_noldus_ajo2alk_cw600/ibi/stancon18_fits/','pubplot'),
                        width = 180, height = 100){
  cowplot::ggsave(filename = file.path(savedir, savename),
                  plot = p, units = 'mm',
                  width = width, height = height)
}
```



# Loading data
```{r variable_naming, include=FALSE}
# Load data, make long format

ibid <- readRDS(params$input_data_rds) %>%
  ungroup() %>% #just to make sure
  gather('variable', 'value', one_of('hr', 'hr.c', 'ibi', 'ibi.c'))
head(ibid)

```

# Descriptive statistics

## Global
```{r}

mystats <- function(ds){
  tibble(n = nrow(ds),
         mean = mean(ds$value),
         median = median(ds$value),
         sd = sd(ds$value),
         mad = mad(ds$value))
}

dstats <- ibid %>%
  group_by(variable) %>%
  do(mystats(.))
dstats
```

## Task
```{r}
dstats_task <- ibid %>%
  group_by(variable, task) %>%
  do(mystats(.))
dstats_task %>% filter(variable == 'hr.c')
```


## Subject
```{r}
dstats_subj <- ibid %>%
  group_by(variable, subject) %>%
  do(mystats(.)) %>%
  ungroup()

tmp <- dstats_subj %>% 
  filter(variable == 'hr') %>%
  mutate(
    mean_c = mean - (dstats %>% filter(variable == 'hr'))$mean,
    median_c = median - (dstats %>% filter(variable == 'hr'))$median) %>%
  select(subject, mean_c, median_c, mean, median, sd, mad)
tmp
```


```{r}
range(tmp$mean_c)
```


## Role 
```{r}
dstats_role <- ibid %>%
  group_by(variable, role) %>%
  do(mystats(.)) %>%
  ungroup()

tmp <- dstats_role %>% 
  filter(variable == 'hr') %>%
  mutate(
    mean_c = mean - (dstats %>% filter(variable == 'hr'))$mean,
    median_c = median - (dstats %>% filter(variable == 'hr'))$median) %>%
  select(role, mean_c, median_c, mean, median, sd, mad)
tmp
```


# Visualizations

## Task
```{r, out.width = '100%'}

p <- ggplot(data = ibid) +
  geom_boxplot(aes(x = task, y = value)) +
  facet_wrap(~ variable, scales = 'free_y', ncol = 2) +
  theme(axis.text.x  = element_text(angle=-90, vjust=0.5, hjust = 0))
p

```

Plot only HR and save for later use:
```{r}
p <- ggplot(data = ibid %>% filter(variable == 'hr')) +
  geom_boxplot(aes(x = task, y = value)) +
  geom_hline(aes(yintercept = 88), linetype = 2) +
  scale_color_brewer(type = 'qual', palette = 'Set2') +
  theme_bw(base_family = 'sans') +
  theme(axis.text.x  = element_text(angle = -45, vjust = 1, hjust = 0),
        plot.margin = grid::unit(c(2,10,2,2), "mm")) +
  labs(x = 'Task', y = 'Hear rate (bpm)',
       title = 'Training data')
p
my_saveplot(p, 'traindata_by-task.pdf', width = 150)

```


## Subject
```{r, out.width = '100%'}

p <- ggplot(data = ibid) +
  geom_boxplot(aes(x = subject, y = value)) +
  facet_wrap(~ variable, scales = 'free_y', ncol = 2) +
  theme(axis.text.x  = element_text(angle=-90, vjust=0.5, hjust = 0))
p

```

```{r}
pd <- ibid %>%
  filter(variable == 'hr') %>%
  mutate(value_c = value - (dstats %>% filter(variable == 'hr'))$mean)
p <- ggplot(data = pd) +
  geom_boxplot(aes(x = sbjnr, y = value_c)) +
  geom_hline(aes(yintercept = 0)) + 
  theme(axis.text.x  = element_text(angle=-90, vjust=0.5, hjust = 0))
p

```

Plot only HR and save for later use:
```{r}
pd <- ibid %>%
  filter(variable == 'hr') %>%
  mutate(value_c = value - (dstats %>% filter(variable == 'hr'))$mean)
p <- ggplot(data = pd) +
  geom_boxplot(aes(x = sbjnr, y = value_c)) +
  geom_hline(aes(yintercept = 0), linetype = 2) +
  scale_color_brewer(type = 'qual', palette = 'Set2') +
  theme_bw(base_family = 'sans') +
  theme(axis.text.x  = element_text(angle = -45, vjust = 1, hjust = 0)) +
  labs(x = 'Subject', y = 'Hear rate (bpm)',
       title = 'Centered training data')
my_saveplot(p, 'traindata_by-subject.pdf', width = 150)

```


## Role
```{r, out.width = '100%'}

p <- ggplot(data = ibid) +
  geom_boxplot(aes(x = role, y = value)) +
  facet_wrap(~ variable, scales = 'free_y', ncol = 2) +
  theme(axis.text.x  = element_text(angle=-90, vjust=0.5, hjust = 0))
p

```



