---
title: "Analysis of IBI/HR hierarchical bayes model"
author: "Jussi Korpela"
date: '2018-08-08'
output:
  html_document:
    theme: united
    toc: yes
  github_document: default
  pdf_document:
    toc: yes
params:
  input_data_rds: /home/jkor/Dropbox/Code/github/hierarchical-bayes-stancon-18/data/ibi-syndata_data.rds
  stan_fit_rds: /home/jkor/Dropbox/Code/github/hierarchical-bayes-stancon-18/results/hr_20180821T143225/ibistan_task_sbj.rds
---

# Info

fit dir: `r dirname(params$stan_fit_rds)`

fit file name: `r basename(params$stan_fit_rds)`

# Rationale

Provides a standard report that can be used to assess IBI hierarchical bayes model convergence and results. Intended to be used to automatically create reports of model fit.

Related to the Stancon 2018 conference abstract.


# Setup
```{r setup, include=FALSE}
require(tidyverse)
require(ggplot2)
require(rstan) # for extractinf stuff from fit
require(bayesplot) # for stan related plots
require(DT) #for showing large tables

# Set general options
knitr::opts_knit$set(echo = TRUE,
                     fig.width = 6,
                     fig.asp = 0.618,
                     out.width = "70%",
                     fig.align = "center")

# todo: temporary fix:
PARAMS <- list(
    use_testing_data = FALSE,
    target_variable = 'hr',
    n_cores = 4,
    chains = 4,
    iterations = 1e4,
    seed = 38809485)
#PARAMS <- attr(fit, 'user_params')

OUT_WIDTH <- 15*100
```

# Loading data and creating descriptive labels
```{r variable_naming, include=FALSE}
# Load data

fit <- readRDS(params$stan_fit_rds)
ibid <- readRDS(params$input_data_rds)


task_labels1 <- ibid %>%
  group_by(task, task_idx) %>%
  count() %>%
  ungroup() %>%
  arrange(task_idx) %>%
  mutate(stan_name = sprintf('w[%d]', task_idx),
         label = paste0(task,'_mean'),
         group = 'task_mean')

subj_labels1 <- ibid %>%
  group_by(sbjnr, sbj_idx) %>%
  count() %>%
  ungroup() %>%
  arrange(sbj_idx) %>%
  mutate(stan_name = sprintf('s[%d]', sbj_idx),
         label = paste0(sbjnr,'_mean'),
         group = 'subj_mean')

role_labels <- ibid %>%
  group_by(role, role_idx) %>%
  count() %>%
  ungroup() %>%
  arrange(role_idx) %>%
  mutate(stan_name = sprintf('r[%d]', role_idx),
         label = paste0(role,'_mean'),
         group = 'role_mean')


task_labels2 <- ibid %>%
  group_by(task, task_idx) %>%
  count() %>%
  ungroup() %>%
  arrange(task_idx) %>%
  mutate(stan_name = sprintf('sd_w_tilde[%d]', task_idx),
         label = paste0(task,'_sd'),
         group = 'task_sd')

subj_labels2 <- ibid %>%
  group_by(sbjnr, sbj_idx) %>%
  count() %>%
  ungroup() %>%
  arrange(sbj_idx) %>%
  mutate(stan_name = sprintf('sd_sbj_tilde[%d]', sbj_idx),
         label = paste0(sbjnr,'_sd'),
         group = 'subj_sd')

misc_labels <- data.frame(
  stan_name = c('beta','sigma_e'),
  label = c('mean_baseline', 'sd_baseline'),
  group = 'baselines'
)


labs <- rbind(as.data.frame(task_labels1 %>%
                              select(stan_name, label, group)),
              as.data.frame(task_labels2 %>%
                              select(stan_name, label, group)),
              as.data.frame(subj_labels1 %>%
                              select(stan_name, label, group)),
              as.data.frame(subj_labels2 %>%
                              select(stan_name, label, group)),
              as.data.frame(role_labels %>%
                              select(stan_name, label, group)),
              misc_labels)
labs <- as.tibble(labs) %>%
  mutate(group = ordered(group,
                         levels = c("baselines",
                                    "task_mean","subj_mean","role_mean",
                                    "task_sd","subj_sd")))
#rownames(labs) <- labs$stan_name

```


# Convergence

It took `r attr(fit, 'computation_time')` `r attr(attr(fit, 'computation_time'),'units')` to fit the model.

## Fit summary
```{r}
fit
```


## Divergences
```{r}
is_divergent <- get_sampler_params(fit,
                                inc_warmup=FALSE)[[1]][,'divergent__']
sum(is_divergent)
```


### Scatterplot with divergences
A plot of one task related effect (`w.1`) with respect to task hierarchy variance `sigma_w`. If the divergences are located in a certain area then that area is obviously hard to sample. If they are scattered randomly, a shorter step size would be appropriate.
```{r}
params_ps <- as.data.frame(rstan::extract(fit, permuted=FALSE))
names(params_ps) <- gsub("chain:1.", "", names(params_ps), fixed = TRUE)
names(params_ps) <- gsub("[", ".", names(params_ps), fixed = TRUE)
names(params_ps) <- gsub("]", "", names(params_ps), fixed = TRUE)
params_ps <- as.tibble(params_ps) %>%
  mutate(is_divergent = ordered(is_divergent,
                                levels = c(1, 0),
                                labels = c('yes','no')))


p <- ggplot(data = params_ps) +
  geom_point(aes(x = w.1, y = log(sigma_w), color = is_divergent)) +
  labs(title = 'Scatterplot projection of the "hierarchy funnel"')
p
```



## Convergence measures as a searchable table
```{r}
diagd <- as.data.frame(summary(fit)$summary)
diagd$parameter <- rownames(diagd)
diagd <- as.tibble(diagd) %>%
  select(parameter, n_eff, Rhat) 
rownames(diagd) <- NULL
diagd <- left_join(diagd, labs, by = c('parameter' = 'stan_name') )
DT::datatable(diagd)
```


## Histograms of n_eff and Rhat

```{r}
p1a <- ggplot(data = diagd) +
  geom_histogram(aes(x = Rhat), binwidth = 0.05)
p1a

p1b <- p1a + facet_wrap(~ group)
p1b


p2a <- ggplot(data = diagd) +
  geom_histogram(aes(x = n_eff), binwidth = 10)
p2a

p2b <- p2a + facet_wrap(~ group)
p2b

# dev stuff:
# pd <- diagd %>%
#   gather(measure, value, n_eff, Rhat)
# pd
# 
# my_plot <- function(ds){
#   p <- ggplot(ds) +
#     geom_dotplot(aes(x = value), method = 'histodot', bins = 100) +
#     scale_y_continuous(NULL, breaks = NULL)
#     #labs(y = unique(ds$measure))
#     #facet_grid(measure ~ ., scales = 'free_x')
#   p
# }
# 
# tmp <- pd %>%
#   group_by(measure) %>%
#   nest() %>%
#   mutate(plot = map(data, my_plot))
# 
# tmp$plot

```


## Convergence of chains: trace plots

```{r convergence-traceplot, out.width = '100%'}

# Plot MCMC trace from a given posterior
# Needs to have three arguments in the order below
my_mcmc_trace <- function(fit, posterior, param_df){
  color_scheme_set("mix-blue-pink")
  p <- mcmc_trace(posterior,
                  pars = param_df$stan_name,
                  n_warmup = attr(fit, 'stan_args')[[1]]$warmup,
                  facet_args = list(labeller = label_parsed))
  p + facet_text(size = 15)
  p
}


# Plot parameters in groups if they exist in posterior
# Uses tidyverse tools for looping over param_df$group
# Outputs a tibble with list-columns
plot_existing <- function(plot_fn, fit, param_df,
                          inc_warmup = F){
  
  # extracted once to save time
  posterior <- rstan::extract(fit, inc_warmup = inc_warmup, permuted = F)
  
  # a wrapper to make map() straightforward
  plotter <- function(param_df){
    plot_fn(fit, posterior, param_df)
  }
  
  tmp <- labs %>% 
    filter(stan_name %in% dimnames(posterior)$parameters) %>%
    group_by(group) %>%
    nest(.key = param_data) %>%
    ungroup() %>%
    mutate(plot = map(param_data, plotter))

  tmp
}

# Generate plots and print
#tmp <- plot_existing(my_mcmc_trace, fit,
# labs %>% filter(group == 'subj_mean'))
tmp <- plot_existing(my_mcmc_trace, fit, labs, inc_warmup = T)
tmp$plot

# dev:
# plot_fn <- function(param_df){
#   my_mcmc_trace(fit, param_df)
# }
# 
# tmp <- labs %>% 
#   group_by(group) %>%
#   nest(.key = param_data) %>%
#   ungroup() %>%
#   filter(group == "task_mean") %>%
#   mutate(plot = map(param_data, plot_fn))
# tmp
# plot_fn(fit, tmp$param_data[[1]])
```



# Visuals

## Posterior distributions of selected variables

```{r posterior-dists,  out.width = '100%'}

# Plot posterior intervals
# Needs to have three arguments in the order below
my_mcmc_intervals <- function(fit, posterior, param_df){

  param_sel <- param_df %>%
    filter(stan_name %in% dimnames(posterior)$parameters)
  posterior_sub <- posterior[ , , param_sel$stan_name]
  dimnames(posterior_sub)[[3]] <- param_sel$label
  
  p <- bayesplot::mcmc_intervals(posterior_sub,
                                 pars = param_sel$label)
  p
}

# Generate plots and print
#tmp <- plot_existing(my_mcmc_intervals, fit,
#  labs %>% filter(group == 'subj_mean'))
tmp <- plot_existing(my_mcmc_intervals, fit, labs, inc_warmup = F)
tmp$plot
# Note: it is important to exclude warmup!

```



```{r posterior_visualization, out.width = OUT_WIDTH, include = F}
# first iteration: too verbose
# 
# posterior <- rstan::extract(fit, inc_warmup = F, permuted = F)
# labs_sel <- labs %>%
#   filter(stan_name %in% dimnames(posterior)$parameters)
# posterior_sub <- posterior[ , , labs_sel$stan_name]
# dimnames(posterior_sub)[[3]] <- labs_sel$label
# 
# 
# my_posterior_plot <- function(posterior_data,
#                               stan_pars,
#                               target_variable){
#   p <- bayesplot::mcmc_intervals(posterior_data,
#                     pars = stan_pars) +
#     labs(x = target_variable)
#   p
# }
# 
# # task mean
# my_posterior_plot(posterior_sub,
#                   as.character(task_labels1$label), PARAMS$target_variable)
# 
# # subject mean
# if ( length(intersect(dimnames(posterior_sub)$parameters,
#                       subj_labels1$label)) > 1 ){
#   my_posterior_plot(posterior_sub,
#                     as.character(subj_labels1$label),
#                     PARAMS$target_variable)                        
#                       }
# 
# # role mean
# if ( length(intersect(dimnames(posterior_sub)$parameters,
#                       role_labels$label)) > 1 ){
#   my_posterior_plot(posterior_sub,
#                     as.character(role_labels$label),
#                     PARAMS$target_variable)
# }
# 
# # task sd
# my_posterior_plot(posterior_sub,
#                   as.character(task_labels2$label), PARAMS$target_variable)
# 
# # subject sd
# if ( length(intersect(dimnames(posterior_sub)$parameters,
#                       subj_labels2$label)) > 1 ){
#   my_posterior_plot(posterior_sub,
#                     as.character(subj_labels2$label),
#                     PARAMS$target_variable)
# }
# 
# # misc
# if ( length(intersect(dimnames(posterior_sub)$parameters,
#                       misc_labels$label)) > 1 ){
#   my_posterior_plot(posterior_sub,
#                     as.character(misc_labels$label),
#                     PARAMS$target_variable)
# }
```

